---
title: "Exercise 14: Moving to Canada"
author: "Michael Manolakis"
date: "April 2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(knitr)
library(leaflet)
options(scipen=9999)

if (Sys.info()['sysname'] == "Windows") {
  if (Sys.info()['nodename'] == "DegreeLaptop" || Sys.getenv('USERNAME') == "Degree Laptop") {
    setwd("C:/Users/Degree Laptop/Desktop/Spring 2025/3 - Econ 256 (Data Vis)/exercise14")}
  else if (Sys.info()['nodename'] == "Michael" || Sys.getenv('USERNAME') == "Michael") {
    setwd("C:/Users/Michael/Desktop/Spring 2025/3 - Econ 256/R Studio Files/Econ256/exercise14")}
} else if (Sys.info()['sysname'] == "Linux") {
  setwd("~/Desktop/R Files/Econ256/exercise14")
} else if (Sys.info()['sysname'] == "Darwin") {
  setwd("~/Desktop/Spring 2025/3 - Econ 256")}

vanhousing <- read_csv("vanhousing.csv")
vantracts <- read_sf(dsn="vantracts", layer="vantracts") %>%
  left_join(
    vanhousing %>%
      group_by(tract) %>%
      summarize(medprice = median(price, na.rm = TRUE)), by = "tract")
```

```{r compile-summary-list, echo=FALSE}
model_beds <- lm(price ~ Beds, data = vanhousing)
model_baths <- lm(price ~ FullBath, data = vanhousing)
model_full <- lm(price ~ Beds + FullBath + west, data = vanhousing)

van_summary <- list(
  correlation_matrix = vanhousing %>%
    select(price, Beds, FullBath, `area_sqft`) %>%
    cor(use = "complete.obs") %>%
    round(2),

  median_price_west_east = vanhousing %>%
    group_by(west) %>%
    summarize(medprice = median(price, na.rm = TRUE)),

  median_price_by_beds = vanhousing %>%
    group_by(Beds) %>%
    summarise(median_price = median(price, na.rm = TRUE)) %>%
    arrange(Beds),

  top10_largest_homes = vanhousing %>%
    arrange(desc(`area_sqft`)) %>%
    slice_head(n = 10) %>%
    select(`area_sqft`, addressshort, price, Beds),

  summary_stats = tibble(
    median_price = median(vanhousing$price, na.rm = TRUE),
    mean_price = mean(vanhousing$price, na.rm = TRUE),
    sd_price = sd(vanhousing$price, na.rm = TRUE),
    median_beds = median(vanhousing$Beds, na.rm = TRUE),
    median_baths = median(vanhousing$FullBath, na.rm = TRUE),
    median_sqft = median(vanhousing$`area_sqft`, na.rm = TRUE)
  ),

  regression_beds = summary(model_beds),
  regression_baths = summary(model_baths),
  regression_full = summary(model_full)
)
```


# Summary of Information and Responses

Appropriate information related to the questions posed for information about Vancouver are found within the following sections, with appropiate titles and details.

The largest home in terms of GLA would be 4766 4th Ave, with 11,950 sqft and a bedroom count of 5, with a price of \$4,132,920. The home with the highest bedroom count would be 2177 30th Ave with a GLA of 3704 sqft, with a price of \$781,900.

The median home price of homes in West Vancouver is \$688,820, while in East Vancouver the median home price is \$633,604.

The table 'Homes that Meet Requirements' shows the homes that meet the criteria of having at least 6 bedrooms, 5 full baths, located in the West side of Vancouver, and priced under $500,000. This is a rare find in Vancouvers housing market, as most homes with such features are typically priced much higher. This is shown fact that there only two homes meet this criteria; 141 41ST Ave, and 163 41ST Ave.

## Homes that Meet Requirements
```{r filtered-homes, echo=FALSE}
filtered_homes <- vanhousing %>%
  filter(Beds >= 6, FullBath >= 5, west == 1, price < 500000) %>%
  select(addressshort, Beds, FullBath, price) %>%
  mutate(price = format(round(price, 0), big.mark = ","))

colnames(filtered_homes) <- c("Address", "Beds", "Full Baths", "Price")
kable(filtered_homes)
```

## Median Price by West/East
```{r median-west-east, echo=FALSE}
west_table <- van_summary$median_price_west_east
west_table$west <- c("East", "West")
colnames(west_table) <- c("", "Median Price")
west_table$`Median Price` <- format(round(west_table$`Median Price`, 0), big.mark = ",")
kable(west_table)
```

## Summary Statistics
```{r summary-stats, echo=FALSE}
summary_table <- van_summary$summary_stats
names(summary_table) <- c("Median Price", "Mean Price", "Standard Deviation", "Median Beds", "Median Baths", "Median Sqft")
summary_table <- summary_table %>%
  mutate(across(everything(), ~ format(round(.x, 0), big.mark = ",")))
kable(summary_table)
```

\newpage
## Median Price by Number of Bedrooms
```{r median-by-bedrooms, echo=FALSE}
beds_table <- van_summary$median_price_by_beds
colnames(beds_table) <- c("Number of Bedrooms", "Median Price")
beds_table$`Median Price` <- format(round(beds_table$`Median Price`, 0), big.mark = ",")
kable(beds_table)
```

## Correlation Matrix
```{r correlation-matrix, echo=FALSE}
cor_table <- van_summary$correlation_matrix
colnames(cor_table) <- rownames(cor_table) <- c("Price", "Beds", "Full Bath", "Area Sqft")
kable(cor_table)
```

## Top 5 Largest Homes by Square Footage
```{r top-5-homes, echo=FALSE}
homes_table <- vanhousing %>%
  arrange(desc(`area_sqft`)) %>%
  slice_head(n = 5) %>%
  select(`area_sqft`, addressshort, price, Beds)
colnames(homes_table) <- c("Area Sqft", "Address", "Price", "Beds")
homes_table$`Area Sqft` <- format(round(homes_table$`Area Sqft`, 0), big.mark = ",")
homes_table$Price <- format(round(homes_table$Price, 0), big.mark = ",")
kable(homes_table)
```

## Top 5 Homes by Bedroom Count
```{r top-5-bedrooms, echo=FALSE}
bedroom_table <- vanhousing %>%
  arrange(desc(Beds)) %>%
  slice_head(n = 5) %>%
  select(Beds, addressshort, price, `area_sqft`)
colnames(bedroom_table) <- c("Beds", "Address", "Price", "Area Sqft")
bedroom_table$Price <- format(round(bedroom_table$Price, 0), big.mark = ",")
bedroom_table$`Area Sqft` <- format(round(bedroom_table$`Area Sqft`, 0), big.mark = ",")
kable(bedroom_table)
```

## Regression Analysis

### Price vs. Number of Bedrooms
```{r regression-beds, echo=FALSE}
kable(as.data.frame(van_summary$regression_beds$coefficients), digits = 3)
```

### Price vs. Number of Bathrooms
```{r regression-baths, echo=FALSE}
kable(as.data.frame(van_summary$regression_baths$coefficients), digits = 3)
```

### Price vs. Bedrooms, Bathrooms, and West/East
```{r regression-full, echo=FALSE}
kable(as.data.frame(van_summary$regression_full$coefficients), digits = 3)
```

### Regression Interpretation

The following regressions estimate the average change in home price associated with adding bedrooms, bathrooms, and living on the west side of Vancouver:
- **Beds Only**: Each additional bedroom is associated with an average increase of **$215,637.70** in home price, without accounting for other features.
- **Full Baths Only**: Each additional full bathroom increases average home price by approximately **$310,628.30**, independent of other characteristics.
- **Combined Model**: When controlling for both bedrooms, bathrooms, and location:
  - An additional bedroom adds **$203,986.40** to home price on average.
  - An additional full bathroom adds **$114,141.30**.
  - Living on the **west side** of Vancouver increases home price by approximately **$594,856.80**, holding other features constant.80, all else held equal.


\newpage
## 
```{r histogram, echo=FALSE, fig.width=6, fig.height=2}
ggplot(data = vanhousing, aes(x = price)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "white") +
  scale_x_continuous(breaks = seq(0, 5000000, by = 1000000),
                     labels = c("0", "$1 Mil", "$2 Mil", "$3 Mil", "$4 Mil", "$5 Mil")) +
  ggtitle("Distribution of House Prices in Vancouver") +
  xlab("Price") +
  ylab("Count")
```

## 
```{r mean-over-time, echo=FALSE, fig.width=6, fig.height=2}
vanhousing %>%
  group_by(year_of_sale) %>%
  summarise(mean_price = mean(price, na.rm = TRUE)) %>%
  ggplot(aes(x = year_of_sale, y = mean_price)) +
  geom_line(color = "darkred", linewidth = 1) +
  scale_x_continuous(breaks = seq(2005, 2015, by = 2)) +
  scale_y_continuous(breaks = seq(600000, 1200000, by = 200000),
                     labels = c("$600K", "$800K", "$1 Mil", "$1.2 Mil")) +
  xlab("Year of Sale") +
  ylab("Mean Price") +
  ggtitle("Change in Mean House Price Over Time")
```

## 
```{r census-tract, echo=FALSE, fig.width=7, fig.height=3.8}
ggplot(data = vantracts) +
  geom_sf(aes(fill = medprice), color = "white", size = 0.1) +
  scale_fill_viridis_c(label = scales::dollar, option = "C", name = "Median Price") +
  labs(title = "Median Home Price by Census Tract in Vancouver") +
  theme_minimal()
```
